\chapter{Poker Without Revealing the Cards}

\label{chap:poker-without-revealing-the-cards}

\begin{abstract}
This chapter studies a dealer-free construction for the dealing phase of Texas
Hold'em among mutually distrustful parties. The objective is not yet to
implement the full game, but to isolate the minimal cryptographic mechanisms
required to instantiate a hand without disclosing private cards before
showdown. We decompose the problem into four stages: deck randomization,
secure multi-party shuffling, secure dealing, and public commitment to private
hands. The current implementation is an instructional demonstrator rather than
a production protocol: it illustrates confidentiality and integrity properties
of the deal while deferring betting, hand evaluation, malicious security, and
network transport.
\end{abstract}

\section{Problem Setting}
Suppose a collection of players wishes to emulate the role of a trusted dealer
in Texas Hold'em. A satisfactory protocol must preserve at least the following
properties:
\begin{enumerate}
    \item \textbf{Card privacy.} No player should learn another player's hole
    cards before the designated reveal phase.
    \item \textbf{Dealing integrity.} No card may be duplicated, omitted, or
    substituted without detection.
    \item \textbf{Transcript consistency.} The participants must agree on a
    single public record of the shuffled deck and the subsequent public cards.
    \item \textbf{Showdown verifiability.} When a player eventually reveals a
    hand, the rest of the table must be able to verify that the hand was fixed
    in advance.
\end{enumerate}

The present chapter restricts attention to the dealing problem. That is, it
addresses how cards are generated, shuffled, privately assigned, and later
committed to, without yet formalizing betting rounds or final winner
determination.

\section{Assumptions and Scope}
The current demonstrator makes three explicit assumptions.
\begin{enumerate}
    \item Each player has access to local randomness.
    \item The players can publish values to a shared transcript visible to all
    participants.
    \item The implementation uses a commutative-encryption toy model in order
    to expose the protocol structure clearly.
\end{enumerate}

The third point is important. The demonstrator is useful as a protocol sketch
and as an executable reference for the staged construction, but it should not
yet be interpreted as a hardened cryptographic system. In particular, no claim
is presently made regarding resistance to malicious deviations, adaptive
adversaries, collusion strategies, or implementation-level side channels.

\section{Four-Stage Construction}
We organize the protocol into four stages so that each requirement can be
examined independently.
\begin{enumerate}
    \item \textbf{Deck randomization.} Establish a hidden permutation of a
    canonical 52-card deck.
    \item \textbf{Secure multi-party shuffling.} Have each player transform the
    deck so that no single player controls the final ordering.
    \item \textbf{Secure dealing.} Deliver hole cards privately and community
    cards publicly.
    \item \textbf{Hand commitment and showdown validation.} Bind each player to
    the dealt hand before revelation.
\end{enumerate}

\section{Stage 1: Deck Randomization}
Let \(D = (c_1,\dots,c_{52})\) denote the canonical deck. The first stage
applies a random permutation to \(D\), thereby producing a deck order that is
not publicly revealed. A public digest of the resulting state is then
published. This digest does not expose the order of the cards, but it fixes a
reference point for the subsequent transcript and prevents ambiguity about
which deck instance is under discussion.

\section{Stage 2: Secure Multi-Party Shuffling}
Each player next applies two transformations to the entire deck: a private
shuffle and a private commutative-encryption layer. Because the encryption
operations commute, the layers may later be removed in different orders
without changing the final plaintext card. This makes it possible to deliver a
private card to one player without disclosing it to the others.

After each player finishes the transformation, a public commitment to the
encrypted deck is published. These intermediate commitments do not reveal the
card values, but they bind the protocol to a single encrypted deck state after
each participant's action.

At the end of this stage, the protocol also publishes a hashed commitment for
each encrypted card that is about to be dealt. Thus, before any decryption
occurs, all participants can see a public digest of the sampled ciphertexts
corresponding to future hole cards and community cards. These digests do not
reveal the dealt cards, but they certify that the later deal came from the
jointly shuffled deck rather than from an after-the-fact substitution.

\section{Stage 3: Secure Dealing}
The dealing phase consumes cards from the jointly shuffled encrypted deck.
Private hole cards are obtained by removing encryption layers in an order that
leaves the final plaintext visible only to the intended recipient. Community
cards are obtained by jointly removing all layers, after which the cards become
public.

At the end of this phase, the transcript should certify that the deal is
internally consistent. In the current demonstrator this is checked by verifying
that no plaintext card identifier appears twice among the dealt hole cards and
community cards. In addition, each drawn encrypted card is checked against the
public hash published in Stage~2, so that the plaintext revealed in this stage
can be traced back to a previously committed random sample from the encrypted
deck.

\section{Stage 4: Commitment to Private Hands}
Once the hole cards have been privately assigned, each player \(P_i\) samples a
private nonce \(N_i\) and publishes a commitment
\[
    C_i = H\!\left(N_i \parallel 2 \parallel h_{i,1} \parallel h_{i,2}\right),
\]
where \(h_{i,1}\) and \(h_{i,2}\) are the player's hole cards and \(H\) is a
cryptographic hash function.

This commitment serves two purposes. First, it binds the player to a specific
private hand before showdown. Second, it does so without disclosing the hand,
provided that the nonce remains secret and the hash function resists inversion
and collisions at the intended security level. At showdown, the player reveals
\(N_i\) together with the private cards, and every participant recomputes
\(C_i\) to validate consistency with the earlier commitment.

\section{Security Interpretation}
The four-stage decomposition isolates the central ideas behind dealer-free
poker:
\begin{enumerate}
    \item secrecy of the final deck order,
    \item joint control over the shuffle,
    \item private delivery of hole cards,
    \item and delayed public verifiability of private hands.
\end{enumerate}

This already captures a substantial portion of the informational structure of
the game. However, it does not yet resolve the full strategic problem of
poker. A complete construction must additionally specify how actions are
recorded, how betting rounds are synchronized, how hand values are computed,
and how disputes are resolved when one or more parties deviate from the stated
protocol.

\section{Implementation Companion}
The executable demonstrator associated with this chapter lives in the source
directory \path{/src/src/poker_without_revealing_the_cards/}. The primary file
is \texttt{poker\_without\_revealing\_the\_cards.c}.
It can be run with:
\begin{verbatim}
cd /src
make -C demostrations demo_poker
\end{verbatim}

The implementation prints stage-by-stage logs, deck commitments, encrypted deck
previews, dealt cards, and showdown validation data. Its purpose is to make the
construction inspectable rather than to claim final protocol completeness.

\section{Limitations and Next Steps}
The current chapter and implementation deliberately stop short of a full poker
system. The following components remain open:
\begin{enumerate}
    \item betting and action privacy,
    \item hand-ranking evaluation,
    \item authenticated network transport,
    \item protection against malicious rather than merely cooperative or
    semi-honest execution,
    \item and stronger proofs for shuffle correctness and commitment validity.
\end{enumerate}

These omissions are not accidental; they define the next research layers. The
current contribution is the disciplined decomposition of the dealing problem
into verifiable stages that can be refined independently.
