\chapter{Tic-Tac-Toe without Revealing the Strategy}
\label{chap:tic-tac-toe-private-strategy}

\begin{quote}
\textbf{Author's Intuition.}

\emph{A perfect strategy in Tic-Tak-Toe is an undefeated one, one that might
win or tie any game. Here is a way to prove to your adversary that you know the
perfect strategy without revealing the strategy. Commit to a session seed;
avoid the adversary cloning the strategy; let the adversary select the initial
conditions; inspect all the Monte Carlo scenarios; investigate whether one can
invert the matrix; and always include code for demonstration.}
\end{quote}

These notes contain several distinct research questions. This chapter
preserves them, but it does not present them as one completed mechanism. A
session identifier, commitment randomness, and a policy-generation seed are
different objects. Random sampling is not an exhaustive proof. No matrix has
yet been defined whose inverse implies anything about Tic-Tac-Toe. These are
open directions, not hidden steps in the construction below.

\begin{center}
\fbox{\begin{minipage}{0.91\textwidth}
\textbf{Current Result: Public Verification by Full Disclosure.}

The current C companion checks a complete, disclosed deterministic policy
through three public representations: a recursive diagnostic checker, a fixed
bottom-up game-DAG evaluator, and deterministic role-specialized Boolean
circuits called \(\mathsf{Core\mbox{-}X}\) and
\(\mathsf{Core\mbox{-}O}\). The circuits check the public game version,
claimant role, empty initial board, complete canonical policy, and fixed-DAG
root value. The policy remains plaintext. ``Witness'' names an input-wire
range; it does not confer secrecy. The circuits contain no
\(\mathsf{Bind}\), commitment, proof system, zero-knowledge mechanism,
authentication, or anti-cloning mechanism.
\end{minipage}}
\end{center}

\begin{abstract}
This chapter separates a private-strategy aspiration from the public predicate
that any later private proof would have to establish. For ordinary
three-by-three Tic-Tac-Toe, we define legal reachable states, a canonical
deterministic positional policy, and a non-loss condition for a fixed player
role. A plain C11 diagnostic checker validates a fully disclosed policy and
exhaustively branches over every legal opponent move from the standard empty
board. A second C11 evaluator expresses the same verdict as a fixed bottom-up
recurrence over the reachable game DAG, without recursion or a
policy-selected successor address. A third dependency-free C11 component emits
and evaluates fixed XOR/AND circuits for the game predicate
\(\mathsf{Core}\). All three remain disclosure baselines, not a solution to
the chapter title. We distinguish
knowledge of a committed policy, policy legality, global non-loss, private
per-state evaluation, and resistance to extraction through repeated queries.
We then state a candidate committed-policy relation for future zero-knowledge
argument-of-knowledge research, compare bounded implementation routes, and
complete the backend-neutral Boolean-circuit step for
\(\mathsf{Core}\). The separate \(\mathsf{Bind}\) predicate, both
commitment layers, and every proof transformation remain open. We identify the
leakage, abort, composition, and setup questions that remain open.
\end{abstract}

\section{Cooperation Problem}

A player may wish to demonstrate that it possesses a strategy that cannot be
defeated while retaining the particular strategy as private information. The
verifier wants evidence stronger than one successful game, but the player does
not want to disclose a complete response table.

The parties are only partially aligned. Both benefit from a precise and
checkable claim. The claimant benefits from keeping its policy private, while
the verifier benefits from finding a losing branch or extracting useful
responses. The narrow research question is:

\begin{quote}
Can a claimant prove knowledge of one fixed, legal, non-losing policy for a
specified role and initial state without revealing that policy beyond the
truth of the claim?
\end{quote}

The first public slice answered a prerequisite question instead: what exact
predicate should be checked if the policy is disclosed? The second fixed its
game-DAG recurrence, and the present slice freezes that recurrence as public
Boolean netlists. None of these steps hides the policy. Defining the predicate
first prevents a future proof system from hiding an ambiguous or incorrect
computation.

\section{Disclosure and Trusted-Evaluator Baselines}

Under full disclosure, the claimant sends the complete strategy to the
verifier. The verifier can check every entry and simulate every opponent
continuation. This provides transparent verification but no policy privacy.

A trusted evaluator offers another baseline. The claimant privately supplies
the policy; the evaluator checks it and announces a verdict. This hides the
policy from the verifier only by moving trust to the evaluator, who learns the
policy and may misreport, retain, or redistribute it.

The public checker in this chapter implements the first baseline. A future
cryptographic construction would seek some of the evaluator's confidentiality
without inheriting its discretionary power. Cryptography is relevant only
after the statement, witness, verifier, and leakage have been fixed.

\section{Fixed Game Model}

The game is ordinary three-by-three Tic-Tac-Toe. Number the cells from $0$
through $8$ in row-major order. Each cell is empty, X, or O. X moves first;
the roles alternate; a row, column, or diagonal of one mark wins; and play
stops immediately after a win. A full board without a winner is a draw. The
initial state $s_0$ is the empty board.

A state is \emph{legal and reachable} only if it occurs along such a play from
$s_0$. In particular, boards with impossible mark counts, simultaneous
winners, or moves after a win are not reachable game states.

Encode cell values as

\[
    \mathsf{EMPTY}=0,\qquad \mathsf{X}=1,\qquad \mathsf{O}=2,
\]

and encode a board $b$ by

\[
    \operatorname{index}(b)=\sum_{i=0}^{8} b_i 3^i.
\]

Thus there are $3^9=19{,}683$ possible byte-addressable board encodings,
although most do not describe legal reachable states. Exhaustive enumeration
from the empty board, without expanding terminal states, gives 5,478 reachable
boards: 2,423 nonterminal X-to-move boards, 2,097 nonterminal O-to-move boards,
and 958 terminal boards. The C tests use these independently expected counts
as finite-model cross-checks, not as cryptographic evidence.

\section{Canonical Deterministic Policy}

Fix a claimant role $p\in\{\mathsf{X},\mathsf{O}\}$. Let $S_p$ be the set
of legal reachable nonterminal states in which $p$ moves, and let $A(s)$
be the empty cells in state $s$. A deterministic positional policy is a
function

\[
    \pi:S_p\longrightarrow\{0,\ldots,8\},
    \qquad \pi(s)\in A(s)\text{ for every }s\in S_p.
\]

Canonical Policy Version~1 represents this function as a 19,683-byte table.
At an index corresponding to a state in $S_p$, the byte is one legal cell
number from $0$ to $8$. Every invalid board, opponent-turn board, terminal
board, and other non-required position contains the distinguished byte
\texttt{0xff}. Consequently, the encoding has no semantically ignored bytes.

The claimant role is a separate public input. The table does not encode the
claimant's identity, authorship, originality, randomness, session, or promise
to follow the policy later.

The original note proposes committing to a session seed. A seed alone commits
to nothing. A binding commitment to that seed can bind the selected policy
only if a public deterministic algorithm

\[
    \pi=\mathsf{GeneratePolicy}(\mathit{seed})
\]

is fixed and that equality is included in the checked relation. Without such
a rule, a seed and a policy are unrelated strings. A unique public session
identifier distinguishes execution context only when it is incorporated into
the committed message, public statement, or authenticated transcript;
commitment randomness is private opening material used by a hiding
construction; and a generation seed may compactly select a policy. None
substitutes for the others.

\section{Legal, Non-Losing, and Minimax-Optimal}

These terms must not be merged.

\begin{itemize}
    \item A \emph{legal} policy selects an empty cell at every required
    claimant-turn state and has the canonical no-move byte everywhere else.
    \item A policy is \emph{non-losing from $s_0$} if, when it is followed
    at every claimant turn and the opponent may make any legal move, every
    terminal branch is a claimant win or a draw.
    \item A policy is \emph{locally minimax-optimal} if its choice has optimal
    game-theoretic value at every state under consideration.
\end{itemize}

For ordinary Tic-Tac-Toe, the value of the empty board under optimal play is a
draw. A non-losing policy therefore realizes the optimal value from $s_0$,
but it need not choose an optimal move at every legal off-path state. The
current checker establishes non-loss only from the fixed empty board. It does
not establish local optimality at every state, uniqueness, or superiority to
another non-losing policy.

Allowing a verifier to choose arbitrary initial conditions changes the claim.
Some legal states are already forced losses for one role. A challenged-state
protocol must define a public allowed set, include the initial state in the
statement, and distinguish a bad continuation from a position whose value was
already losing. This remains open.

\section{Public Exhaustive Predicate}

For a role $p$ and canonical policy $\pi$, define
$\mathsf{PublicNonLose}(p,\pi)=1$ exactly when:

\begin{enumerate}
    \item every required entry is present and selects an empty cell;
    \item every non-required entry is \texttt{0xff}; and
    \item from $s_0$, following $\pi$ at claimant turns and branching over
    every legal opponent move never reaches an opponent win.
\end{enumerate}

The verifier first classifies the complete legal reachable-state graph. It
rejects missing, out-of-range, occupied-cell, and unexpected entries. It then
performs a deterministic depth-first traversal. At a claimant turn it follows
the policy's single move. At an opponent turn it explores every empty cell in
ascending order. It stops at every win or full-board draw. On the first
opponent win it returns the complete chronological move sequence.

This is exhaustive enumeration, not Monte Carlo sampling. A random sample can
miss the sole losing branch of an otherwise successful policy, and therefore
cannot by itself establish the universal no-loss predicate. Sampling could
support a separately stated statistical test under a named distribution and
error bound, but that would be a different claim.

The original matrix-inversion note is also not part of this predicate. No
matrix, encoding, or theorem currently connects invertibility to legal play or
non-loss. If such a connection is discovered, it should first be stated as a
separate mathematical relation and compared with direct exhaustive checking.

\section{C Implementation Companion and Evidence}

\begin{sloppypar}
The preserved public-checker interface is \path{include/ac/ttt.h}; both
evaluators remain in \path{src/protocols/ttt.c}; and their tests remain in
\path{tests/test_ttt.c}. The Boolean interface is split between
\path{include/ac/bool_circuit.h} and
\path{include/ac/ttt_core_circuit.h}; its implementations are
\path{src/circuits/bool_circuit.c} and
\path{src/circuits/ttt_core_circuit.c}; and its tests are
\path{tests/test_bool_circuit.c} and
\path{tests/test_ttt_core_circuit.c}. The demonstration is
\path{src/tic_tac_toe_without_revealing_the_strategy/ttt_demo.c}. The protected
checker interface, implementation, and original 13-test suite remain
byte-identical independent baselines. The implementation uses C11,
caller-owned storage, and the standard library only. It performs no heap
allocation, randomness, hashing, commitment, network communication, or
libsodium call.
\end{sloppypar}

Both evaluators distinguish API execution from the mathematical verdict. Null
pointers and invalid roles are API errors. A malformed but readable policy is
a successfully computed invalid-policy verdict. A canonical policy with an
opponent-winning continuation is a losing verdict accompanied by a replayable
trace from the recursive checker. The fixed-DAG report deliberately contains
no invented trace or DFS exploration count. The public reference-policy
generator uses deterministic backward
induction; it is a convenience for demonstrations and tests, not a secret or
cryptographic mechanism.

The test suite checks both roles, an independent graph and history census,
malformed encodings, late losing branches, deterministic repeated reports,
terminal-state handling, every legal one-entry alternative of both generated
reference policies, and a deterministic set of multi-entry policies. The
single-entry sweep contains 6,208 X policies and 5,439 O policies. It is
exhaustive only at Hamming distance one from those two reference tables, not
over all canonical policies. Such tests establish enumerated implementation
behaviour for tested builds. The induction argument below establishes the
mathematical recurrence; neither it nor the tests are machine-checked formal
verification of the C program. They do not establish policy privacy, zero
knowledge, knowledge extraction, authorship, or future behaviour.

Most importantly, the checker receives the entire 19,683-byte table. It does
not implement the title's non-revelation goal.

\section{Five Claims That Must Remain Separate}

\subsection{Knowledge of a Strategy}

A commitment-shaped value does not demonstrate that its sender knows an
opening. Formal proof of knowledge requires an extractor-based definition
relating verifier acceptance to recovery of a witness
\cite{bellare1993definingknowledge}. The current checker has no commitment and
no extractor.

\subsection{Legality}

Once a complete policy is disclosed, legality is an ordinary deterministic
property. An unopened commitment cannot be scanned for missing, occupied, or
unexpected moves. Merely committing to bytes proves none of their semantics.

\subsection{Non-Loss}

Once a legal policy is disclosed, finite exhaustive traversal establishes the
implemented no-loss predicate. One successful game, a set of favourable
examples, or an unspecified Monte Carlo loop does not cover every adversarial
branch.

\subsection{Private Per-State Evaluation}

Returning $\pi(s)$ necessarily reveals that output. If the state and output
are public, a future claimant might return the move with a zero-knowledge proof
that it is consistent with the same committed policy. If the verifier's query
state must also be hidden, the problem becomes private two-party computation.
Neither mechanism is present here.

\subsection{Resistance to Cloning}

A reusable interface that returns the exact deterministic move on arbitrary
states exposes the table one entry at a time. Cryptography cannot conceal an
answer that the interface deliberately releases. Query restrictions,
statefulness, fresh session policies, or randomized policies may change how
quickly useful information accumulates, but each requires its own leakage
definition.

\section{Future Committed-Policy Relation}

Commitment schemes formalize hiding and binding properties under stated
assumptions; Naor's construction is one foundational reference
\cite{naor1991bitcommitment}. That work does not prove the repository's
instructional hash commitment or the Tic-Tac-Toe predicate.

A future protocol may use the public instance

\[
 x=(v_G,p,s_0,C,\mathit{ctx})
\]

and the private witness

\[
 w=(\pi,r),
\]

where $v_G$ fixes the game and policy versions, $s_0$ fixes the initial state,
$C$ is a policy commitment, $\mathit{ctx}$ fixes the protocol and session
context, and $r$ is its opening information. Define the relation

\[
 \begin{aligned}
 R(x,w)=1 \quad\Longleftrightarrow\quad{}
 &v_G=1 \;\land\; p\in\{\mathsf{X},\mathsf{O}\}
     \;\land\; s_0=\mathsf{EmptyBoard} \\
 &{}\land\;\mathsf{Open}\bigl(C,\mathit{ctx},
     \mathsf{Encode}(v_G,p,s_0,\pi),r\bigr)=1 \\
 &{}\land\;\mathsf{PublicNonLose}(p,\pi)=1.
 \end{aligned}
\]

Zero knowledge asks that the verifier learn no more from a proof than what is
already implied by the public statement. The foundational definition is due
to Goldwasser, Micali, and Rackoff
\cite{goldwasser1989knowledgecomplexity}. Goldreich, Micali, and Wigderson show
the conceptual breadth of zero-knowledge proofs for NP under their stated
assumptions \cite{goldreich1991proofsvalidity}. This existence result does not
select a practical backend, prove knowledge automatically, or establish the
security of a concrete composition.

The author's full technical goal would require an instantiated protocol having
both a zero-knowledge property and a proof- or argument-of-knowledge property
for $R$, under named setup, adversary, commitment, channel, and composition
assumptions. The computational target selected below is an argument of
knowledge. This chapter implements none of that machinery. In particular, the
existing instructional hash commitment is not silently promoted into a proof
system.

\section{Selecting a Private-Proof Route}

The immediate target is one global zero-knowledge argument of knowledge for the
relation $R$, not a service that returns policy moves. The public instance and
accept/reject outcome remain visible. A later instantiation that meets the
selected zero-knowledge definition against a malicious verifier should reveal
no additional information about the witness beyond that public instance and
the truth of the statement. A per-state interface instead returns $\pi(s)$ and
therefore reveals one policy output on every query. Private per-state
evaluation is a separate two-party-computation problem with its own input
privacy, output leakage, fairness, and repeated-query requirements.

This target is provisional, not achieved. It concerns one standalone
interactive execution by classical probabilistic polynomial-time algorithms:
completeness is perfect or explicitly bounded; computational knowledge
soundness is stated against a malicious prover through an extractor for the
complete witness; and computational zero knowledge is stated against a
malicious verifier. For the combinatorial soundness calculation, the honest
verifier samples independent uniform challenges after the first proof-internal
commitments. The malicious-verifier zero-knowledge target must instead handle
arbitrary probabilistic polynomial-time verifier challenge strategies under the
selected transformation. The security definitions concern the direct abstract
prover--verifier transcript. Human identity, transport authentication, relays,
delivery, timing, concurrency, resets, composition, endpoint compromise,
fairness, and recovery are outside this initial target.

The relation also has two costs that must remain visible. Write

\[
 R(x,w)=\mathsf{Core}(v_G,p,s_0,\pi)
        \land \mathsf{Bind}(C,\mathit{ctx},v_G,p,s_0,\pi,r),
\]

The predicate $\mathsf{Core}$ checks the game version, valid role, and empty
initial state. It also evaluates $\mathsf{PublicNonLose}$, which already
includes complete canonical-policy validation. The predicate $\mathsf{Bind}$
checks the exact commitment opening and context. Proving only $\mathsf{Core}$
does not connect the policy to $C$. Proving only $\mathsf{Bind}$ says nothing
about legality or non-loss.
Dropping $C$ entirely would merely prove knowledge of some non-losing ordinary
Tic-Tac-Toe policy, a capability that is already public and easy to reproduce.

\subsection{Bounded Comparison}

The foundational generic result of Goldreich, Micali, and Wigderson shows why
an NP relation can be approached through zero knowledge
\cite{goldreich1991proofsvalidity}. A generic reduction remains valuable
theory, but implementing such a reduction as this chapter's companion would
add substantial representation and communication machinery without teaching
the structure of the finite-game predicate.

The MPC-in-the-head method of Ishai, Kushilevitz, Ostrovsky, and Sahai converts
a suitable multiparty computation into a zero-knowledge protocol
\cite{ishai2007zeroknowledgemp}. ZKBoo develops a concrete version for Boolean
circuits using symmetric primitives \cite{giacomelli2016zkboo}. Proposition
4.2 of ZKBoo analyzes its basic three-message protocol in the
commitment-hybrid model, assuming a correct, 2-private $(2,3)$-decomposition of
the circuit computation. In that model it is a Sigma protocol with special
honest-verifier zero knowledge and 3-special soundness. These properties do not
automatically provide malicious-verifier zero knowledge.

The proof-internal commitments to simulated views are distinct from the public
policy commitment $C$ checked by $\mathsf{Bind}$. A concrete experiment must
separately justify the binding and hiding of the view commitments, any
pseudorandom-tape assumptions, their encodings and domains, and their transcript
cost. Naming ZKBoo makes neither commitment layer compatible with the current
instructional hash commitment.

For one idealized repetition, a uniformly sampled challenge in
$\{1,2,3\}$ leaves the combinatorial soundness term at most $2/3$. With
independent repetitions and independently uniform challenges chosen after the
first messages, that term is $(2/3)^r$; the smallest integer making only this
term at most $2^{-128}$ is

\[
 r=\left\lceil\frac{128}{\log_2(3/2)}\right\rceil=219.
\]

This equation is not a total soundness bound or proof-size estimate for our
relation. A concrete argument must also include the binding-failure term of its
chosen view commitment. The Boolean cost of \(\mathsf{Core}\) is now measured
for the frozen construction below. The costs of \(\mathsf{Bind}\),
proof-internal commitments, simulated views, transformations, and transcripts
remain unmeasured. A \(\mathsf{Core}\) gate count is not a proof-size
estimate. Transformations beyond the basic Sigma protocol are required for the chapter's
malicious-verifier zero-knowledge and knowledge target. Making the protocol
non-interactive with Fiat--Shamir would additionally enter the random-oracle
model and require exact transcript binding and challenge sampling. This chapter
selects neither transformation yet.

Transparent succinct systems offer a different tradeoff. Spartan, for
example, provides argument-of-knowledge constructions for R1CS without a
relation-specific trusted setup in its transparent variants
\cite{setty2020spartan}. Its stack includes an arithmetic constraint system,
sum-check, and both computation and polynomial commitments. This is a credible
later production-oriented comparison point, but importing or reimplementing
such machinery would violate the present objective of a small, inspectable
next experiment.

Accordingly, MPC in the head is the leading family for a later explicitly
educational interactive experiment, conditional on circuit measurement and a
separate analysis of the full zero-knowledge and knowledge transformations.
It is not an implemented protocol, a selected production backend, or a current
security claim. Generic NP reductions and succinct arithmetic systems remain
comparison points; a trusted evaluator and private per-state computation remain
separate baselines.

\subsection{Completed Implementation Gate: A Fixed Game DAG}

The second executable evaluator still contains no proof system. It expresses
the public predicate as a fixed bottom-up directed acyclic graph and keeps the
recursive checker as an independent diagnostic baseline.

Independent finite enumeration gives 5,478 reachable states, of which 4,520
are nonterminal and 958 are terminal. The terminal states comprise 626 X wins,
316 O wins, and 16 draws. There are 8,631 legal edges out of X-turn states and
7,536 out of O-turn states, or 16,167 public state edges in total. By contrast,
the unmerged legal history tree has 549,946 nodes. The merged state graph is
therefore the natural fixed computation; these counts do not estimate a
cryptographic proof size. Independent tests also count 255,168 terminal legal
games in that unmerged tree.

The base-three board encoding supplies a particularly small topological
schedule. If cell $a$ is empty and the public side to move has numeric encoding
$t\in\{1,2\}$, then

\[
  \mathsf{child}(i,a)=i+t3^a>i.
\]

Consequently, one ascending scan of all 19,683 encodings can propagate
reachability and classify the graph, and one descending scan can evaluate
every child before its parent. A complete ascending scan between them validates
every policy byte. The implementation needs no recursion, queue, heap, edge
table, or stored ply schedule.

Canonical validation is part of the relation rather than a precondition hidden
from it. At every required state the evaluator compares the policy byte with
each public cell number. It never reads a board or successor at an address
chosen by that byte. It scans the full table, records the first malformed entry
only for diagnostics, continues the fixed evaluation, and gives the
invalid-policy verdict precedence over the root value. This prevents a missing
or nonsensical move from being misreported merely as a losing policy.

For every reachable state $i$, let $q_i$ denote whether the claimant avoids
loss from that state while following the policy. The recurrence is
well-founded from later plies to earlier plies; the implementation's
descending index scan is an equivalent reverse topological order because every
legal child index exceeds its parent index:

\[
q_i=
\begin{cases}
1, & i\text{ is terminal and is not an opponent win},\\
0, & i\text{ is an opponent win},\\
\displaystyle\bigvee_{a\in A(i)}
 \left([\pi(i)=a]\land q_{\mathsf{child}(i,a)}\right),
    & i\text{ is a claimant-turn state},\\
\displaystyle\bigwedge_{a\in A(i)}q_{\mathsf{child}(i,a)},
    & i\text{ is an opponent-turn state}.
\end{cases}
\]

Here $[\pi(i)=a]$ is a Boolean equality bit. Every state, legal move $a$, and
successor index is public; only the equality result and derived values depend
on the policy. Canonical-policy validation ensures that exactly one legal move
is selected at each claimant state. The $q_i$ values are deterministically
constrained intermediates, not independent semantic witness inputs; a backend
may nevertheless encode them as auxiliary wires or witness entries. The
recurrence accepts exactly when the empty-board value is one.

\noindent\textbf{Proposition.}
\emph{DAG and recursive predicate equivalence.}
For either claimant role and every Canonical Policy Version~1 table $\pi$, the
fixed-DAG root value equals the Boolean verdict of the recursive checker.

\textbf{Proof.}
For a legal reachable state $s$, let $F(s)$ be the recursive semantic value
after canonical validation: it is zero exactly for an opponent-win terminal;
at a claimant turn it is the value of the single policy child; and at an
opponent turn it is the conjunction of every legal child value. We prove
$q(s)=F(s)$ by strong induction on the number of empty cells.

At a terminal state the two definitions agree directly. At a nonterminal
claimant state, canonical validity gives one legal move
$a^*=\pi(s)$. Every selector equality except $[\pi(s)=a^*]$ is zero, so the
disjunction reduces to $q(\mathsf{child}(s,a^*))$; the induction hypothesis
makes this $F(\mathsf{child}(s,a^*))=F(s)$. At a nonterminal opponent state,
both definitions are the conjunction of the values of all legal children, and
the induction hypothesis applies to each child. Therefore $q(s)=F(s)$ at every
reachable state and in particular at the empty board. $\square$

Malformed-table agreement is a separate validation statement. Both
implementations use the same required states, reason precedence, and first
ascending-index error. The differential tests compare that evidence, but do
not compare the recursive checker's first losing path or history-node counts
with a DAG report.

For claimant X, the C recurrence performs 8,631 policy-to-cell equalities,
8,631 selector term conjunctions, 8,631 selector disjunction folds, and 7,536
opponent conjunction folds. For claimant O, 7,536 replaces each selector
count and there are 8,631 opponent folds. Both roles scan 19,683 graph slots,
19,683 policy bytes, and 19,683 value slots; derive 5,478 reachable values; and
visit 16,167 public edges. These are operations of this C evaluator, not gates
in an optimized Boolean circuit.

The explicit scratch payload is two 19,683-byte arrays, one for state kinds and
one for values, plus a nine-byte board: 39,375 bytes in total. The serialized
policy remains exactly 19,683 bytes. This count excludes scalar locals, the
caller-owned report, compiler padding, and the call frame, so it is not a
whole-process memory measurement.

The differential corpus includes every legal Hamming-distance-one alternative
of the two generated reference policies: 6,208 for X and 5,439 for O. It also
includes 256 deterministically generated multi-entry policies per role and
malformed representatives. This does not enumerate the full canonical spaces,
whose sizes are

\[
  9\,7^{72}5^{756}3^{1372}
  \quad\text{for X},\qquad
  8^9 6^{252}4^{1140}2^{696}
  \quad\text{for O}.
\]

The multi-entry cases are a reproducible corpus, not a random sample, and no
probability bound is claimed. Tests can reveal implementation disagreement;
the induction gives the mathematical reason the two recurrences agree and may
still fail to expose a shared low-level C defect.

This equivalence gate supplies the semantics for the Boolean construction
below. The commitment computation remains separate rather than disappearing
inside the game predicate. Proof-internal view commitments would be an
additional cryptographic cost and assumption, not part of
\(\mathsf{Bind}\).

\subsection{Boolean Circuits for Core}

The implementation emits two circuits, \(\mathsf{Core\mbox{-}X}\) and
\(\mathsf{Core\mbox{-}O}\). Each specialization still receives and checks
the public role byte. The public input is the four-byte sequence

\begin{center}
\texttt{[version][role][initial-index-high][initial-index-low]}.
\end{center}

The index is an unsigned big-endian integer. Version~1 accepts only its
specialized role and initial index zero. The policy input is the unchanged
19,683-byte Canonical Policy Version~1 table. The IR calls that range
\emph{witness input}, but the evaluator receives every byte in plaintext.
Within each packed byte, bit zero is assigned first.

Wire zero is constant false and wire one is constant true. Public bits come
next, followed by policy bits and then gate outputs. Every gate has two earlier
wire operands and is either XOR or AND. Circuit shape, operands, and output wire
depend on the public game and role specialization, not on policy contents.
Required policy bytes reuse their legal-move equality wires between canonical
validation and the claimant recurrence; non-required bytes are compared with
\texttt{0xff}. The circuit follows the same public successor locations and
descending board-index recurrence as the fixed-DAG evaluator.

The canonical serialization is byte-defined rather than a dump of C structs.
It has a 32-byte header, big-endian integer fields, and one nine-byte record per
gate, for exactly \(32+9G\) bytes. Construction, evaluation, depth analysis,
and serialization use caller-owned storage. Evaluation uses one byte per wire
and applies a best-effort volatile wipe before returning. That wipe is not a
guarantee about registers, compiler-created copies, operating-system state, or
the caller-owned public and policy buffers.

\begin{center}
\begin{tabular}{l@{\hspace{1.5em}}r@{\hspace{2em}}r}
\hline
Quantity & \(\mathsf{Core\mbox{-}X}\) & \(\mathsf{Core\mbox{-}O}\) \\
\hline
Public input bits & 32 & 32 \\
Policy/witness input bits & 157,464 & 157,464 \\
AND gates & 215,022 & 209,313 \\
XOR gates & 31,350 & 27,344 \\
Total gates & 246,372 & 236,657 \\
Total wires & 403,870 & 394,155 \\
Output depth & 31 & 30 \\
Gate storage (bytes) & 2,956,464 & 2,839,884 \\
Evaluator scratch (bytes) & 403,870 & 394,155 \\
Canonical serialization (bytes) & 2,217,380 & 2,129,945 \\
\hline
\end{tabular}
\end{center}

These are exact counts for this frozen construction, confirmed by separately
analyzing and parsing the emitted circuits. They are not lower bounds,
optimized backend costs, proof sizes, transcript sizes, running-time results,
or security levels. The named builder scratch occupies 255,888 bytes on the
tested C11 ABI; this records \texttt{sizeof} for that layout rather than an
ABI-independent whole-process memory claim.

The bounded differential corpus covers the two reference policies, known
canonical losing policies, all 256 values at policy byte zero, named malformed
classes, 256 stratified legal one-entry mutations and 128 deterministic
canonical multi-entry policies per role, and invalid public version, role, and
initial-index inputs. Representative cases use the scalar production
evaluator. A separate test-only 64-lane interpreter reads the canonical bytes
independently so many policies can be compared with both preserved C
evaluators without changing the public API. This corpus is not exhaustive over
the enormous canonical policy spaces, and testing is not formal verification.

Serialization is not a commitment. Evaluating \(\mathsf{Core}\) does not
evaluate the complete relation
\(R=\mathsf{Core}\land\mathsf{Bind}\). The external policy commitment,
proof-internal commitments, soundness, extraction, zero knowledge, privacy,
interaction, and production security remain absent.

\section{Repeated-Query Reconstruction}

\textbf{Proposition (exact finite-oracle reconstruction).}
Let $D$ be a finite domain and let $\pi:D\rightarrow A$ be deterministic.
Any party permitted to query an oracle $O_\pi(s)=\pi(s)$ at every $s\in D$
can reconstruct $\pi$ using $|D|$ distinct queries.

\textbf{Proof.}
Enumerate the elements of $D$. Query each element once and store the returned
pair $(s,O_\pi(s))$. The resulting table agrees with $\pi$ at every member
of its domain and is therefore the same function. No computational assumption
is involved. $\square$

For Canonical Policy Version~1, unrestricted access over all claimant-turn
states reconstructs every required entry. If access is limited to games
actually reached from $s_0$, the proposition does not imply recovery of
unqueried off-path entries, even though those entries are legally reachable;
nevertheless, every observed move reveals the policy at one visited state.
Repeated adaptive games expose an expanding partial table.

A randomized policy changes exact recovery into statistical estimation:
repeated queries reveal samples from its output distribution. An anti-cloning
claim would then need a query budget, a target approximation metric, and a
definition of what counts as a successful clone.

One global zero-knowledge argument for the non-loss relation avoids serving a
per-state policy oracle. It can certify a property without offering the
strategy as a service. It also does not prevent the verifier from independently
constructing another non-losing Tic-Tac-Toe policy.

\section{Attacks, Abort, Leakage, and Composition}

The public checker is designed to expose malformed and losing policies rather
than conceal them. Relevant cases include missing entries, out-of-range moves,
occupied-cell moves, stray bytes on opponent or terminal states, role
confusion, post-win boards, and a single losing line hidden behind many safe
branches.

Future private protocols add different risks:

\begin{itemize}
    \item a commitment may be replayed in a different game, role, or session
    if its context is incomplete;
    \item a malicious party may publish a syntactically valid value for which
    it knows no opening;
    \item a biased challenge or sampled checker may avoid the losing branch;
    \item selected openings or policy responses accumulate information across
    sessions;
    \item an interactive prover may abort before the verifier accepts;
    \item a recipient may learn a per-state output and then abort before
    providing its own promised action; and
    \item a protocol satisfying a stated zero-knowledge or knowledge-soundness
    property in a stand-alone execution may not retain that property under
    concurrent, reset, or adaptively interleaved executions.
\end{itemize}

No protocol can force a silent participant to send the next message. An abort
establishes non-completion, not motive, guilt, or network cause. Fairness,
guaranteed output delivery, availability, authentication, and recovery are
therefore not consequences of the future relation $R$. They would require
separate mechanisms and claims.

\section{Ethical and Institutional Interpretation}

The attractive general idea is broader than this small game: a party may wish
to demonstrate that a private policy satisfies a public safety condition
without surrendering every private preference. The separation between policy
and proof can protect legitimate autonomy.

It can also hide power. A proof establishes only the encoded predicate. It
does not show that the predicate is fair, that excluded states were chosen
justly, that affected people can appeal, or that the hidden policy lacks other
harmful behaviour. In a consequential institution, the authority to define
the state space, allowed initial conditions, objective, proof circuit, and
update rules matters at least as much as witness privacy.

Tic-Tac-Toe is intentionally harmless and finite. It is a laboratory for this
distinction, not evidence that the same construction should govern people.

\section{What Cryptography Does Not Solve}

Even a correct future zero-knowledge argument of knowledge for $R$ would not:

\begin{itemize}
    \item decide whether non-loss is the right objective;
    \item establish authorship, originality, identity, or exclusive knowledge;
    \item prove that the claimant will follow the policy after the proof;
    \item prevent independent discovery of another non-losing policy;
    \item hide outputs intentionally released by a policy oracle;
    \item force completion, explain an abort, or guarantee availability;
    \item make arbitrary losing initial conditions winnable; or
    \item give matrix inversion a role that has not been mathematically defined.
\end{itemize}

Zero knowledge can constrain what a transcript reveals about a witness. It
cannot choose a just predicate or turn a public, easily computed capability
into a unique strategic advantage.

\section{Limitations and Open Problems}

The current result is limited to fixed ordinary three-by-three Tic-Tac-Toe, a
deterministic positional policy, one fixed role, and the standard empty board.
The complete policy is disclosed in memory. The implementation contains no
protocol-message parser, persistence, identity, transport, commitment, proof
system, private computation, random policy, or generalized-game analysis.

The following questions remain open:

\begin{enumerate}
    \item Which public deterministic generation rule, if any, should connect
    the author's proposed session seed to a policy?
    \item Which initial positions may a verifier choose, and how should their
    pre-existing minimax values alter the claim?
    \item Which separately versioned \(\mathsf{Bind}\) circuit should connect
    the complete canonical policy to an external commitment and context, under
    which commitment assumptions, and at what measured cost?
    \item Which concrete MPC-in-the-head transformation can prove knowledge of
    the complete committed witness while meeting a malicious-verifier
    zero-knowledge definition; under which commitment, setup, and composition
    assumptions; and with what rounds, transcript, and proof size?
    \item Should per-state evaluation reveal the query, the move, both, or only
    a derived game action?
    \item What query budget and leakage metric would make a meaningful
    anti-cloning claim?
    \item Does the author's matrix intuition correspond to a useful linear or
    algebraic representation, or should it remain a metaphor?
    \item Can a role-specific compact internal witness be expanded
    canonically to the complete Version~1 policy without changing the message
    bound by the commitment?
\end{enumerate}

These are research questions. The recursive checker, fixed-DAG evaluator, and
Boolean Core remain public-disclosure baselines against which later answers
must be compared.

\section{Exercises}

\begin{enumerate}
    \item Independently enumerate the legal reachable states and reproduce the
    four state counts in this chapter.
    \item Construct a canonical policy containing exactly one reachable losing
    branch, and explain why random play may miss it.
    \item Give an example showing why non-loss from $s_0$ does not imply
    locally optimal moves at every off-path state.
    \item Define a public policy-generation algorithm and extend $R$ so that
    a private seed, rather than a policy table, is the witness. List the new
    assumptions this introduces.
    \item Prove the finite-oracle reconstruction proposition for the role you
    choose, and state why it does not automatically reconstruct unqueried
    off-path states when access is restricted to actual games.
    \item Choose a legal state that is already losing for the player to move.
    Explain why unrestricted verifier-selected initial conditions make the
    original non-loss claim impossible.
    \item Describe the difference between a zero-knowledge proof for $R$ and
    a zero-knowledge proof of knowledge for $R$. Identify what an extractor
    must output in the latter case.
    \item Prove directly from the base-three board encoding that every legal
    child index exceeds its parent index. Reimplement the bottom-up recurrence
    without calling either C evaluator and compare all three verdicts on a
    precisely described finite corpus.
    \item Independently parse, recount, and evaluate both serialized Core
    circuits. Explain why agreement on a bounded corpus proves neither policy
    privacy nor circuit equivalence for every possible policy.
\end{enumerate}
